%       Copyright (c) 2001-2014 by INRIA and SDTools, All Rights Reserved.
%       Use under OpenFEM trademark.html license and LGPL.txt library license
%       $Revision: 1.21 $  $Date: 2018/10/12 17:07:45 $


%-----------------------------------------------------------------------------
\rtop{iimouse}{iimouse}


\noindent Mouse related callbacks for GUI figures.

\rsyntax\begin{verbatim}
iimouse
iimouse('ModeName')
iimouse('ModeName',Handle)
\end{verbatim}\nlvs

\rmain{Description}


The non current {\sl SDT 3} version of this function is included in OpenFEM. Use the {\tt help} \ts{fecom} command to get help.


% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\rmain{See also}

\noindent  \iicom, \fecom, \iiplot, \iiplot

\input{lsutil.tex}

%------------------------------------------------------------------------------
\rtop{nopo}{nopo}

Imports nopo files (cf. Modulef)

\rsyntax\begin{verbatim}
 model      = nopo('read -v -p type FileName')
 [Node,Elt] = nopo('read -v -p type FileName')
\end{verbatim}\nlvs

\rmain{Description}
 
\ruic{nopo}{read}{}
 The \ts{-v} option is used for verbose output. The optional \ts{-p} \tsi{type} gives the type of problem described in the nopo file, this allows proper translation to OpenFEM element names.  Supported types are

 {\tt '2D','3D','AXI','FOURIER','INCOMPRESSIBLE','PLAQUE','COQUE'}.

\rmain{See also}  \medit


%------------------------------------------------------------------------------
\rtop{medit}{medit}

Export to {\tt Medit} format

\rsyntax\begin{verbatim}
[indnum,scale] = medit('write FileName',model)
[indnum,scale] = medit('write FileName',model,def,[opt])
[indnum,scale] = medit('write FileName',model,def,'a',[opt])
[indnum,scale] = medit('write FileName',model,[],strain)
[indnum,scale] = medit('write FileName',model,def,strain,[opt])
[indnum,scale] = medit('write FileName',model,def,strain,'a',[opt])
\end{verbatim}\nlvs


\rmain{Description}

{\tt Medit} is an interactive mesh visualization software, developed by the Gamma project at INRIA-Rocquencourt. 

{\tt Medit} executable is freely available at \href{http://www-rocq.inria.fr/gamma/medit}{\tt http://www-rocq.inria.fr/gamma/medit}.
Documentation can also be obtained.

{\tt medit} is an interface to {\tt Medit} software. This function creates files needed by {\tt Medit} and runs the execution of these files in {\tt Medit}. Users must download and install {\tt Medit} themselves. It is not provided in {\tt OpenFEM}.

Input arguments are the following :

\ruic{medit}{FileName:}{}
file name where information for Medit will be written, no extension must be given in FileName.
\ruic{medit}{model}{ :}
a structure defining the model. It must contain at least fields .Node and .Elt.
\ruic{medit}{def}{ :}
a structure defining deformations that users want to visualize. It must contain at least fields .def and .DOF.
\ruic{medit}{strain}{ :}
structure defining coloring, must at least contain :

* fields .data and .EltId  if coloring depends on elements

* fields .data and .DOF    if coloring depends on nodes

Strain can be obtained by a call to {\tt fe\_stres}.\\
For example, {\tt strain=fe\_stres('ener',FEnode,FEel0,pl,il,md1,mdof)} generates a structure with fields .data and .EltId (depending on elements), and {\tt strain = fe\_stres('stress mises',FEnode,FEel0,pl,[],def,mdof)}  generates a structure with fields .data and .DOF (depending on nodes). See demo {\tt d\_ubeam} or test {\tt test\_medit}.
\ruic{medit}{opt}{ :}
option vector, {\tt opt = [numdef nb\_imag scale\_user]} with 

* {\tt numdef} : mode to display number

* {\tt nb\_imag} : number of files to create the animation of deformations

* {\tt scale\_user}  : display scale (a parameter for increasing the deformations) 

\ruic{medit}{indnum}{ :}
returns the nodes numbering used by Medit 

\ruic{medit}{scale}{ :}
returns the scale that was used to display the deformations


\rmain{Use}

\begin{verbatim}
medit('write FileName',model)
%displays the mesh defined by model
medit('write FileName',model,def,[opt])
%displays the mesh defined by model in a window and
% the deformation defined by def in an other window
medit('write FileName',model,def,'a',[opt])
%animates deformations defined by def on model
medit('write FileName',model,[],strain)
%displays the mesh defined by model and colors it with the help of strain
medit('write FileName',model,def,strain,[opt])
%displays the mesh defined by model in a window and 
%the deformation defined by def with colors due to strain in an other window
medit('write FileName',model,def,strain,'a',[opt])
%animates the deformations defined by def on model and
% colors them with the help of strain
\end{verbatim}

Note that nodes and faces references are given to Medit. You can visualize faces references by pressing the right mouse button, selecting ``Shading'' in the ``Render mode'' menu and then selecting ``toggle matcolors'' in the ``Colors, Materials'' menu.

\input{of2vtk.tex}

%------------------------------------------------------------------------------
\rtop{ofutil}{ofutil}


OpenFEM utilities

\rsyntax\begin{verbatim}
 ofutil commands
\end{verbatim}\nlvs

\rmain{Description}

This function is used for compilations, path checking, documentation generation, ...

 Accepted commands are
 
\lvs\begin{tabular}{@{}p{.15\textwidth}@{}p{.85\textwidth}@{}}
{\tt   Path     } & checks path consistency with possible removal of SDT\\
{\tt   mexall   } & compiles all needed DLL\\
{\tt   of\_mk    } & compiles of\_mk.f (see openfem/mex directory)\\
{\tt   nopo2sd  } & compiles nopo2sd.c (located in openfem/mex directory)\\
{\tt   sp\_util  } & compiles sp\_ufil.c\\
{\tt   zip      } & creates a zip archive of the OpenFEM library \\
{\tt   hevea    } & generates documentation with HEVEA\\
{\tt   latex    } & generates documentation with LaTeX\\
%
\end{tabular}
 
%  HP-UX start MATAB R13 : setenv LD_PRELOAD /usr/lib/libF90.sl;matlab

%------------------------------------------------------------------------------
\input{ofact.tex}

%------------------------------------------------------------------------------
\input{sp_util_stack.tex}
%------------------------------------------------------------------------------






